\section{Model}

\subsection{Distributed System} \label{sec:distsys}

A distributed system is a set of interconnected machines that follow a set of instructions to reach a common goal.
The set of instructions is called protocol.
The machines are called network nodes or parties.
The connections between the machines are called network.
The nodes may use the network to exchange information.

\subsection{Time} \label{sec:time}

We model continuous time.
Our default time variable is $t$, the default unit is seconds.
If not stated otherwise, protocol execution starts at time $t = 0$.

\subsection{Information DAG} \label{sec:infodag}

We organize information in a directed acyclic graph (DAG) that grows over time.
Each vertex represents one datum.
Nodes can extend the DAG by appending a new datum $x$ to a set $\parents(x)$ of existing data.
We say that datum $a$ links to datum $b$, or that $a$ was appended to $b$, if and only if $b \in \parents(a)$.

Let $x_1, \dots, x_n$ be a sequence of data.
We say that datum $x_n$ inherits from datum $x_1$, if and only if
$x_{i-1} \in \parents(x_{i})$ for all $i = 2, \dots, n$.

The protocol may impose restrictions on what data can be appended when.
E.\,g., in Bitcoin, data are called blocks.
Each appended block must have exactly one parent, must inherit from the genesis block, and must satisfy a proof-of-work criterion.
By virtue of the first restriction, the DAG becomes a tree.
The nodes use Nakamoto's consensus protocol to agree on and extend upon the same branch of this tree.

\subsection{Partial Visibility} \label{sec:visibility}

Not all nodes see all information at all times.
Right after creation, a datum is only visible to the node that appended the data.
When a datum becomes visible to a node, it stays visible in the future.
Data becomes visible according to the communication rules.

\subsection{Communication} \label{sec:communication}

A node can share visible data with other nodes.
We assume a $\Delta$-synchronous network model.
I.\,e., if a node shares datum $x$ at time $t$, then $x$ is visible to all other nodes at time $t + \Delta$.
